---
title: "Wine quality EDA"
author: ""
date: "today"
# date: "`r Sys.Date()`"
output:
  html_document:
    code_folding: hide
    number_sections: false
    toc: yes
    toc_depth: 3
    toc_float: yes
  pdf_document:
    toc: yes
    toc_depth: '3'
---

```{r init, include=FALSE}
# some of common options (and the defaults) are: 
# include=T, eval=T, echo=T, results='hide'/'asis'/'markup',..., collapse=F, warning=T, message=T, error=T, cache=T, fig.width=6, fig.height=4, fig.dim=c(6,4) #inches, fig.align='left'/'center','right', 
library(ezids)
# knitr::opts_chunk$set(warning = F, results = "markup", message = F)
knitr::opts_chunk$set(warning = F, results = "hide", message = F)
options(scientific=T, digits = 3) 
# options(scipen=9, digits = 3) 
# ‘scipen’: integer. A penalty to be applied when deciding to print numeric values in fixed or exponential notation.  Positive values bias towards fixed and negative towards scientific notation: fixed notation will be preferred unless it is more than ‘scipen’ digits wider.
# use scipen=999 to prevent scientific notation at all times
```

## Wine EDA

**Import the dataset into R**  
There are 1599 observations of red-wine and 4898 observations of white-wine
```{r}
wine_red <- data.frame(read.csv("./dataset/winequality-red.csv",sep = ";")) 
wine_white <- data.frame(read.csv("./dataset/winequality-white.csv",sep = ";")) 
summary(wine_red)
summary(wine_white)
str(wine_red)
str(wine_white)
#wine = read.csv("/Users/nischaysaicherukuri/Downloads/wine-quality-white-and-red.csv")

``` 

Combine the two datasets into one
```{r}
wine_red['category'] = "red"
wine_white['category'] = "white"
wine <- rbind(wine_red,wine_white)

```
```{r}
head(wine)
```


Take a look if there is null and na values
```{r}
colSums(is.na(wine))
```

## EDA Analysis

First,let's take a look of the quality distrubution between white and red wine.
```{r result = TRUE}
unique(wine$quality)
table(wine$quality)
table(wine_white$quality)
library(ggplot2)
ggplot(wine, aes(x=quality, color=category)) +
  geom_histogram(fill="white", position="dodge")+
  scale_x_continuous(limits = c(3, 9), breaks = seq(3, 9, 1)) +
  theme(legend.position="top")+
  xlab('Quality of Wine') +
  ylab('Number of Wines')
```
In the given data set, wine scores are in range [3,8]. The majority of wines  have either 5 or 6 ranking for their quality. The overall distribution is similar for both white and red wine.

Secondly, we would like to see how each independent variables differ from white and red wine.
1. fixed.acidity
```{r result = TRUE}
ggplot(wine, aes(x=fixed.acidity, color=category)) +
  geom_boxplot()
```
2. volatile.acidity
```{r result = TRUE}
ggplot(wine, aes(x=volatile.acidity, color=category)) +
  geom_boxplot()
```
3.citric.acid
```{r result = TRUE}
ggplot(wine, aes(x=citric.acid, color=category)) +
  geom_boxplot()
```
4.residual.sugar
```{r result = TRUE}
ggplot(wine, aes(x=residual.sugar, color=category)) +
  geom_boxplot()
```
5.chlorides
```{r result = TRUE}
ggplot(wine, aes(x=chlorides, color=category)) +
  geom_boxplot()
```
6.free.sulfur.dioxide
```{r result = TRUE}
ggplot(wine, aes(x=free.sulfur.dioxide, color=category)) +
  geom_boxplot()
```
7.total.sulfur.dioxide
```{r result = TRUE}
ggplot(wine, aes(x=total.sulfur.dioxide, color=category)) +
  geom_boxplot()
```
8.density
```{r result = TRUE}
ggplot(wine, aes(x=density, color=category)) +
  geom_boxplot()
```
9.pH
```{r result = TRUE}
ggplot(wine, aes(x=pH, color=category)) +
  geom_boxplot()
```
10.sulphates
```{r result = TRUE}
ggplot(wine, aes(x=sulphates, color=category)) +
  geom_boxplot()
```
11.alcohol
```{r result = TRUE}
ggplot(wine, aes(x=alcohol, color=category)) +
  geom_boxplot()
```
From the boxplots, we can see that the variation is differed significantly in following variables:
fixed.acidity
residual.sugar
total.sulfur.oxide
free.sulfur.dioxide
chlorides
volatile.acidity


## Testing

Now, let's take a look on correlation between each variable 
```{r result = TRUE,warning=F}
loadPkg("corrplot")
#wine_noca_red = subset(wine_red, select = -c(category) )
#wine_noca_redcor = cor(wine_noca_red )
#corrplot(wine_noca_redcor,type="upper")

#wine_noca_white = subset(wine_white, select = -c(category) )
#wine_noca_whitecor = cor(wine_noca_white)
#corrplot(wine_noca_whitecor,type="upper")

wine_noca = subset(wine, select = -c(category) )
wine_nocacor = cor(wine_noca )
corrplot(wine_nocacor,type="upper")
```
As we can see from the correlation diagram, the below variables are mostly corrleated to the quality. And we would like to dig further to see the effect
quality vs alcohol
quality vs density
quality vs volatile.acidity
quality vs chlorides

quality vs alcohol
quality vs sulphates
quality vs citric.acid
quality vs volatile.acidity

We would anaysis in follwing steps:
1. normality test - qq plot & histagram
2. correlation test 
3. scatter plot&boxplot- check the relationship
4. Annova test - compare means of our attribute across the wines and check if differences are statistically significant compared to the quality of a wine.

1.quality vs alcohol
The distribution of alcohol data is right skewed
```{r result = TRUE,warning=F}
qqnorm(wine$alcohol, pch = 1, frame = FALSE)
qqline(wine$alcohol, col = "steelblue", lwd = 2)
ggplot(wine, aes(x=alcohol)) + geom_histogram()
cor.test(wine$alcohol,wine$quality)
```

```{r result = TRUE,warning=F}
ggplot(wine, aes(x=alcohol, y=quality, color=category, shape = category)) +
  geom_point() + 
  geom_smooth(method=lm, se=FALSE, fullrange=TRUE)+
  ggtitle("alcohol vs Quality")

```
```{r result = TRUE,warning=F}
wine_cat <- wine
wine_cat$quality <- factor(wine_cat$quality)
ggplot(wine_cat, aes(x=quality, y=alcohol, fill=category)) + 
    geom_boxplot() +
    facet_wrap(~category)+
    ggtitle("alcohol vs Quality")

anova_alcohol = aov(alcohol ~ quality, data=wine_cat)
summary(anova_alcohol)
```

2.quality vs density
Density plot looks normally distributed
```{r result = TRUE,warning=F}
qqnorm(wine$density, pch = 1, frame = FALSE)
qqline(wine$density, col = "steelblue", lwd = 2)
ggplot(wine, aes(x=density)) + geom_histogram()
cor.test(wine$density,wine$quality)
```
```{r result = TRUE,warning=F}
ggplot(wine, aes(x=density, y=quality, color=category, shape = category)) +
  geom_point() + 
  geom_smooth(method=lm, se=FALSE, fullrange=TRUE)+
  ggtitle("density vs Quality")

```
```{r result = TRUE,warning=F}
ggplot(wine_cat, aes(x=quality, y=density, fill=category)) + 
    geom_boxplot() +
    facet_wrap(~category)+
    ggtitle("density vs Quality")

anova_density = aov(density ~ quality, data=wine_cat)
summary(anova_density)
```


3.quality vs volatile.acidity
The distribution of volatile.acidity data is almost normal,However, there is a small tail on the right side of the plot
```{r result = TRUE,warning=F}
qqnorm(wine$volatile.acidity, pch = 1, frame = FALSE)
qqline(wine$volatile.acidity, col = "steelblue", lwd = 2)
ggplot(wine, aes(x=volatile.acidity)) + geom_histogram()
cor.test(wine$volatile.acidity,wine$quality)
```

```{r result = TRUE,warning=F}
ggplot(wine, aes(x=volatile.acidity, y=quality, color=category, shape = category)) +
  geom_point() + 
  geom_smooth(method=lm, se=FALSE, fullrange=TRUE)+
  ggtitle("volatile.acidity vs Quality")
```
```{r result = TRUE,warning=F}
ggplot(wine_cat, aes(x=quality, y=volatile.acidity, fill=category)) + 
    geom_boxplot() +
    facet_wrap(~category)+
    ggtitle("volatile.acidity vs Quality")

anova_volatile.acidity = aov(volatile.acidity ~ quality, data=wine_cat)
summary(anova_volatile.acidity)
```

4.quality vs chlorides
The distribution looks like normally distributed but is also right skewed 
```{r result = TRUE,warning=F}
qqnorm(wine$density, pch = 1, frame = FALSE)
qqline(wine$density, col = "steelblue", lwd = 2)
ggplot(wine, aes(x=density)) + geom_histogram()
cor.test(wine$density,wine$quality)
```

```{r result = TRUE,warning=F}
ggplot(wine, aes(x=chlorides,y=quality, color=category, shape = category)) +
  geom_point() + 
  geom_smooth(method=lm, se=FALSE, fullrange=TRUE)+
  ggtitle("chlorides vs Quality")
```
```{r result = TRUE,warning=F}
ggplot(wine_cat, aes(x=quality, y=chlorides, fill=category)) + 
    geom_boxplot() +
    facet_wrap(~category)+
    ggtitle("chlorides vs Quality")

anova_chlorides = aov(chlorides ~ quality, data=wine_cat)
summary(anova_chlorides)
```






